Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

 While the introduction of new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally important for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality. Conventionally, the methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account. For example, the Continuous Ranked Probability Score (CRPS) can be used to evaluate the accuracy of probabilistic predictions. It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function. These measures all fail to take sufficient account of whether the right direction of change is being taken, which can often be highly relevant in practice. The information about whether an estimator (be it a measurement device, forecast or nowcas) can accurately predict the increase or decrease is important to determine whether a statement can be made about the direction of development of the actual value on the basis of the estimator's prediction results. In the following, we first provide an overview of the characteristics and evalutations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.

Measurement aims to obtain accurate and reliable data about the current state of a system. Often a certain parameter or variable is measured at regular intervals over a certain period of time. To detect the systematic difference between one measurement method with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}). In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}). 
Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}).

One important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction, when the gold standard detects changes during the measurement process. 
Ideally, the signs of the changes between two consecutive points in time (with a defined time interval) measured by the new method and the gold standard method should always be the same. 
In this case, we call the new method trending perfect in line with the gold Standard method and refer to this type of agreement as trendability. 
This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature (\cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}).

@@

The evaluation of trending is also interesting for the assessment of nowcasts.
In contrast to conventional forecasting methods, Nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields, such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}). In economics, Nowcasting can be used, for example, to predict important statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
Nowcasting methods use indicators that are related to the target variable but can be collected at a higher frequency and published in a timely manner to produce and update early estimates for the target variable during the relevant period (\cite{castle2017forecasting}). For example, in the context of forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024, before the official statistics are published.
In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}). 
The nowcasting method can be used, for example, to correct the daily Covid case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}). 
In this context, nowcasting is an estimation procedure in which the value of a target variable (e.g. from today) is estimated on the basis of current preliminary measurements, which is only finalised at a later point in time. 
At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.  A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.

like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge. However, appropriate measures for assessing trending ability should be constructed, because a nowcast with strong trending ability provides crucial information about the development of the target variable (for example, the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.

@@

\subsection*{Contributions.} 
The main contributions of this paper are summarized below
\begin{itemize}

>>>>

Measurement, prediction and forecasting methods are becoming increasingly important and sophisticated in a world of fast developing technologies and increasing data availability. While the introduction of new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally important for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality. Conventionally, the methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account. For example, the Continuous Ranked Probability Score (CRPS) can be used to evaluate the accuracy of probabilistic predictions. It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function. These measures all fail to take sufficient account of whether the right direction of change is being taken, which can often be highly relevant in practice. The information about whether an estimator (be it a measurement device, forecast or nowcas) can accurately predict the increase or decrease is important to determine whether a statement can be made about the direction of development of the actual value on the basis of the estimator's prediction results. In the following, we first provide an overview of the characteristics and evalutations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.

Measurement aims to obtain accurate and reliable data about the current state of a system. Often a certain parameter or variable is measured at regular intervals over a certain period of time. In addition to the use of a paired t-test to determine systematic differences (\cite{watson2010method}) between two measurement series, various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}). Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}). As already described above, one important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction, when the gold standard detects changes during the measurement process. 
Ideally, the signs of the changes between two consecutive points in time (with a defined time interval) measured by the new method and the gold standard method should always be the same. 
In this case, we call the new method trending perfect in line with the gold Standard method and refer to this type of agreement as trendability. 
This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature (\cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}).

@@

The evaluation of trending is also interesting for the assessment of nowcasts.
In contrast to conventional forecasting methods, Nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields, such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}). In economics, Nowcasting can be used, for example, to predict important statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
Nowcasting methods use indicators that are related to the target variable but can be collected at a higher frequency and published in a timely manner to produce and update early estimates for the target variable during the relevant period (\cite{castle2017forecasting}). For example, in the context of forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024, before the official statistics are published.
In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}). 
The nowcasting method can be used, for example, to correct the daily Covid case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}). 
In this context, nowcasting is an estimation procedure in which the value of a target variable (e.g. from today) is estimated on the basis of current preliminary measurements, which is only finalised at a later point in time. 
At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.  A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change. Like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge. However, appropriate measures for assessing trending ability should be constructed, because a nowcast with strong trending ability provides crucial information about the development of the target variable (for example, the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.

@@

\textbf{Contributions.} The main contributions of this paper are summarized below
\begin{itemize}

### 2024-03-07 sections/03-trending.tex
Update on Overleaf.

Let $F_{t | \tau} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $\tau$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
For trending assessment, we compare the probability of positive change with the occurrence of positive changes. 
The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value. \unsure{Andere Formulierung für: one minus?}
As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
For nowcasts, the probability is calculated by
\begin{equation*}
    p_t = 1 - F_{t+l | t+l} (y_{t})\quad t = 1, \dots, T-l;
\end{equation*}
while for forecasts, it is given by
\begin{equation*}
    p_t = 1 - F_{t+l | t} (y_{t})\quad t = 1, \dots, T-l.
\end{equation*}
Let $z_t$ denote the indicator that the observed change at time $t$ is positive, that is,
\begin{equation*}
    z_t = \ind{\diffyt > 0} \quad t = 1, \dots, T-l.
\end{equation*}
The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=1}^{T-l}$ and the observed changes $\mathbf{z} = (z_t)_{t=1}^{T-l}$.
The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
Dichotomous forecasts predict a binary outcome, for example, a positive or negative change.
Numerical measures for forecast quality are scoring rules.

>>>>

Let $F_{t | \tau} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $\tau$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
For trending assessment, we compare the probability of positive change with the occurrence of positive changes. 
The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value. \unsure{Andere Formulierung für: one minus?}
\unsure{vielleicht sowas: The probability $p_t$ of a positive change is given by the probability that the predicted value at the forecasting time is greater than the known value at the issure time. }
As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
For nowcasts, the probability is calculated by
\begin{equation*}
    p_t = 1 - F_{t+l | t+l} (y_{t})\quad t = 1, \dots, T-l;
\end{equation*}
while for forecasts, it is given by
\begin{equation*}
    p_t = 1 - F_{t+l | t} (y_{t})\quad t = 1, \dots, T-l.
\end{equation*}
Let $z_t$ denote the indicator that the observed change at time $t$ is positive, that is,
\begin{equation*}
    z_t = \ind{\diffyt > 0} \quad t = 1, \dots, T-l.
\end{equation*}
The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=1}^{T-l}$ and the observed changes $\mathbf{z} = (z_t)_{t=1}^{T-l}$.
The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
Dichotomous forecasts predict a binary outcome, for example, a positive or negative change.
Numerical measures for forecast quality are scoring rules.
